Revisited hidden weight bit function

### 2024-10-29 RevisedHiddenWeight.tex
part 4.3 with comments.

\begin{property}[Cryptographic properties of functions from $\WD{1}{n}$ \label{prop:wwd1}]
Let $n\in \N^*$, the following hold for functions in $\WD{1}{n}$.
\begin{itemize}
	\item Nonlinearity of majority functions.
	\item Nonlinearity of HWBF.
	\item Nonlinearity bound WWA.
\end{itemize}

\end{property}

@@

	We prove that the maximum of $\left|-z^2+2z+1\right|$ on $\dis$ is $2\sqrt 2$ by showing that the maximum of $\left|-z^2+2z+1\right|^2$ on $\dis$ equals $8$.
    %It remains to prove that the maximum of $\left|-z^2+2z+1\right|^2$ on $\dis$ equals $8$. 
    Using that $|w|^2=w\overline w$ for every complex number $w$, we obtain for all $z\in \dis$:
    \[
        \left|-z^2+2z+1\right|^2=\left(z^2-2z-1\right)\left(\overline z^2-2\overline z-1\right)=6-\left(z^2+\overline z^2\right)=6-2\mathrm{Re}\left(z^2\right),
    \]
    where $\mathrm{Re}\left(z^2\right)$ denotes the real part of $z^2$. Since the minimum of $\mathrm{Re}\left(z^2\right)$ on $|z|=1$ is $-1$, we conclude that the maximum of $\left|-z^2+2z+1\right|^2$ on $\dis$ is $6-2(-1)=8$.

@@

\subsection{Generalization to a family of weightwise quadratic functions}\label{sec:general}

\begin{definition}
	Let $n\in \N$ be an even number and $t \in \N $ no greater than $n/2$, we denote by $d_{t,n}$ the function given by: $d_{t,n}(x)=\sum_{i=1}^t x_{2i-1} x_{2i}$.

@@

In particular, $d_{n/2,n}(x)=d_n(x)$.

We focus on the Walsh spectrum restricted to the slices of any function $d_{t,n}$

@@

	\begin{align*}
	\wtk{d_{t,n}}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{d_{t,n}(x)+ax}=\sum_{x \in \Ekn{k}{n}} (-1)^{d_{t,t}(y)+by +cz}\\
	&=\sum_{\ell=0}^{2t} \sum_{\substack{y \in \Ekn{\ell}{2t}\\z \in \Ekn{k-\ell}{n-2t}}} (-1)^{d_{t,t}(y)+by +cz}\\
	&=\sum_{\ell=0}^{2t} \sum_{y \in \Ekn{\ell}{2t}} (-1)^{d_{t,t}(y)+by} \left( \sum_{z \in \Ekn{k-\ell}{n-2t}} (-1)^{cz}\right)\\
	&=\sum_{\ell=0}^{2t} \Dnka{2t}{\ell}{b} \kraw{k-\ell}{\w(c)}{n-2t}.\qedhere
	\end{align*}	
\end{proof}

@@

Consequently, we get a generalization of Theorem \ref{theorem:bound_D_nka}:

@@

\begin{proof}
    By applying Property \ref{property:cauchy_estimate}, we get that a bound on $|\wtk{d_{t,n}}{k}(a)|$ is given by the maximum $M'$ on the circle $|z|=1$ of $|P_b(z)|\cdot|(1+z)^u\cdot(1-z)^v|$. As previously established, $|P_b(z)|$ is bounded on $|z|=1$ by $(2\sqrt 2)^t=2^{3t/2}$, and $|(1+z)^u\cdot(1-z)^v|$ is bounded on $|z|=1$ by $2^{u+v}=2^{n-2t}$. Therefore, we conclude that $M'$ is bounded by $2^{3t/2}2^{n-2t}=2^{n-t/2}$, and hence this is also a bound for $|\wtk{d_{t,n}}{k}(a)|$. 
\end{proof}

\begin{Corollary}
    Let $t, n\in \N$ with $n$ even and $t\le n/2$. For every $a\in\F_2^n$, the following holds:
    \[
        |\wt{d_{t,n}}(a)|\leq 1+n\cdot 2^{n-t/2}.
    \]
\end{Corollary}

@@

When one is given an explicit value for $t$ in terms of $n$, one can be more careful when employing the proof strategy of Theorem~\ref{theorem:bound_D_nka_generalized}. Let us for instance consider the case $t=n/3$:

\begin{proposition}\label{proposition:walsh_t_over_3}
    Let $n\in\N$ be divisible by $6$. Then for every $a\in\F_2^n$ and every $k\in[0,n]$, it holds that:
    \[
        |\wtk{d_{n/3,n}}{k}(a)|\leq2^{3n/4},\quad|\wt{d_{n/3,n}}(a)|\leq 1+n\cdot 2^{3n/4}.
    \]
\end{proposition}

\begin{proof}
    As before, we want to bound the maximum of $|P_{a,t}(z)|$ on $|z|=1$ for $a\in\F_2^n$, and this will then be an upper bound on $|\wtk{d_{t,n}}{k}(a)|$ for every $k\in[0,n]$. Write $a=(b,c)$ with $b\in\F_2^{2t}$ and let $(p,q,r)=(p(b),q(b),r(b))$, and further consider $u=n-2t-\w(c)$ and $v=\w(c)$ as in Definition~\ref{defi:P_a_t}. Since we want to find a bound of $|P_{a,t}(z)|$ on $|z|=1$ which works for all $a\in\F_2^n$, it makes sense to only consider those $a$ which are actually candidates for maximizing $|P_{a,t}(z)|$, namely those for which only one among $(p,q,r)$ is non-zero, and for which only one among $(u,v)$ is non-zero. This leaves us with the following cases to consider for $(p,q,r;u,v)$:
    \begin{gather*}
        (t,0,0;n-2t,0),\quad(0,t,0;n-2t,0),\quad(0,0,t;n-2t,0),\\
         (t,0,0;0,n-2t),\quad(0,t,0;0,n-2t),\quad(0,0,t;0,n-2t).
    \end{gather*}
    Since swapping both $(p,q)$ and $(u,v)$ has the same effect as replacing $z$ by $-z$ in $P_{a,t}(z)$ and therefore does not change the maximum of $|P_{a,t}(z)|$ on $|z|=1$, we can further restrict ourselves to the following tuples for $(p,q,r;u,v)$:
    \[
        (t,0,0;n-2t,0),\quad(0,t,0;n-2t,0),\quad(0,0,t;n-2t,0).
    \]
    Let us for instance find an upper bound of $|P_{a,t}(z)|$ in case $(p,q,r;u,v)=(t,0,0;n-2t,0)$. Since $n-2t=t$ by assumption, we get:
    \[
        P_{a,t}(z)=[(-z^2+2z+1)(1+z)]^t.
    \]
    Writing $z=\cos\alpha+\mathrm i\sin\alpha$, we obtain:
    \[
        |P_{a,t}(z)|^2=[(6-2\cos2\alpha)(2+2\cos\alpha)]^t.
    \]
    Through calculus, one finds that $|(6-2\cos2\alpha)(2+2\cos\alpha)|<2^{9/2}$, and hence that $|P_{a,t}(z)|<2^{9t/4}=2^{3n/4}$. The same upper bound works for the remaining two candidates for $(p,q,r;u,v)$, and we can therefore conclude that $|\wtk{d_{n/3,n}}{k}(a)|\leq 2^{3n/4}$ and that $|\wt{d_{n/3,n}}(a)|\leq 1+n\cdot 2^{3n/4}$.
\end{proof}

\begin{remark}
    The proof of Proposition~\ref{proposition:walsh_t_over_3} can be redone for other values of $t$. In Figure~\ref{fig:walsh_bounds_generalization}, we provide bounds of the form $\max_{a\in\F_2^n}|\wt{d_{\alpha n/40,n}}(a)|\leq 1+n\cdot 2^{\beta n}$, where $\alpha\in[0,20]$ is represented on the $x$-axis and $\beta$ is represented on the $y$-axis. 
\end{remark}

\begin{figure}
	\centering
	\includegraphics[width=9cm]{walsh_bounds_generalization.pdf}
	\caption{For each point $(\alpha,\beta)$, we have $\max_{a\in\F_2^n}|\wt{d_{\alpha n/40,n}}(a)|\leq 1+n\cdot 2^{\beta n}$.}
    \label{fig:walsh_bounds_generalization}
\end{figure}

>>>>

\begin{Prop}[Cryptographic properties of functions from $\WD{1}{n}$ \label{prop:wwd1}]
Let $n\in \N^*$, the following hold for functions in $\WD{1}{n}$.
\begin{itemize}
	\item Nonlinearity of majority functions.
	\item Nonlinearity of HWBF.
	\item Nonlinearity bound WWA.
\end{itemize}

\end{Prop}

@@

	We prove that the maximum of $\left|-z^2+2z+1\right|$ on $\dis$ is $2\sqrt 2$ by showing that the maximum of $\left|-z^2+2z+1\right|^2$ on $\dis$ equals $8$.
    %It remains to prove that the maximum of $\left|-z^2+2z+1\right|^2$ on $\dis$ equals $8$. 
    Using that $|w|^2=w\overline w$ for every complex number $w$, we obtain for all $z\in \dis$:
    \[
        \left|-z^2+2z+1\right|^2=\left(z^2-2z-1\right)\left(\overline z^2-2\overline z-1\right)=6-\left(z^2+\overline z^2\right)=6-2\mathrm{Re}\left(z^2\right),
    \]
    where $\mathrm{Re}\left(z^2\right)$ denotes the real part of $z^2$. Since the minimum of $\mathrm{Re}\left(z^2\right)$ on $\dis$ is $-1$, we conclude that the maximum of $\left|-z^2+2z+1\right|^2$ on $\dis$ is $6-2(-1)=8$.

@@

\subsection{Generalization to a family of weightwise quadratic functions}\label{sec:general}
In this part, we generalize the results of Section~\ref{sec:Cauchy} to all weightwise quadratic function such that for all $k\in [1,n]$ $f_k$ has exactly $t$ quadratic terms with no shared variables.

\begin{definition}
	Let $n\in \N^*$ be an even number and $t \in \N $ no greater than $n/2$, we denote by $d_{t,n}$ the function given by: $d_{t,n}(x)=\sum_{i=1}^t x_{2i-1} x_{2i}$.

@@

In particular, $d_{n/2,n}(x)=d_n(x)$. 
We focus on the Walsh spectrum restricted to the slices of any function $d_{t,n}$:

@@

%	\begin{align*}
%	\wtk{d_{t,n}}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{d_{t,n}(x)+ax}=\sum_{x \in \Ekn{k}{n}} (-1)^{d_{t,t}(y)+by +cz}\\
%	&=\sum_{\ell=0}^{2t} \sum_{\substack{y \in \Ekn{\ell}{2t}\\z \in \Ekn{k-\ell}{n-2t}}} (-1)^{d_{t,t}(y)+by +cz}\\
%	&=\sum_{\ell=0}^{2t} \sum_{y \in \Ekn{\ell}{2t}} (-1)^{d_{t,t}(y)+by} \left( \sum_{z \in \Ekn{k-\ell}{n-2t}} (-1)^{cz}\right)\\
%	&=\sum_{\ell=0}^{2t} \Dnka{2t}{\ell}{b} \kraw{k-\ell}{\w(c)}{n-2t}.%\qedhere
%	\end{align*}

	\begin{align*}
\wtk{d_{t,n}}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{d_{t,n}(x)+a \cdot x}=\sum_{x \in \Ekn{k}{n}} (-1)^{d_{t,2t}(y)+b \cdot y +c \cdot z}=\sum_{\ell=0}^{2t} \sum_{\substack{y \in \Ekn{\ell}{2t}\\z \in \Ekn{k-\ell}{n-2t}}} (-1)^{d_{t,2t}(y)+b \cdot y +c \cdot z}\\
&=\sum_{\ell=0}^{2t} \sum_{y \in \Ekn{\ell}{2t}} (-1)^{d_{t,2t}(y)+b \cdot y} \left( \sum_{z \in \Ekn{k-\ell}{n-2t}} (-1)^{c \cdot z}\right)=\sum_{\ell=0}^{2t} \Dnka{2t}{\ell}{b} \kraw{k-\ell}{\w(c)}{n-2t}.%\qedhere
\end{align*}

@@

Consequently, we get a generalization of Theorem~\ref{theorem:bound_D_nka}:

@@

\begin{proof}
    By applying Property \ref{property:cauchy_estimate}, we get that a bound on $|\wtk{d_{t,n}}{k}(a)|$ is given by the maximum $M'$ on the circle $\dis$ of $|P_b(z)|\cdot|(1+z)^u\cdot(1-z)^v|$. 
    As previously established in Theorem~\ref{theorem:bound_D_nka}, $|P_b(z)|$ is bounded on $\dis$ by $(2\sqrt 2)^t=2^{3t/2}$, and $|(1+z)^u\cdot(1-z)^v|$ is bounded on $\dis$ by $2^{u+v}=2^{n-2t}$. 
    Therefore, we conclude that $M'$ is bounded by $2^{3t/2}2^{n-2t}=2^{n-t/2}$, and hence this is also a bound for $|\wtk{d_{t,n}}{k}(a)|$. 
\end{proof}

\begin{Corollary}\label{cor:gen}
    Let $t, n\in \N$ with $n$ even and $t\le n/2$. For every $a\in\F_2^n$, the following holds:
    \[
        |\wt{d_{t,n}}(a)|\leq 1+n\cdot 2^{n-t/2}.
    \]
\end{Corollary}

@@

When a specific value of $t$ is given in terms of $n$, the proof strategy for Theorem~\ref{theorem:bound_D_nka_generalized} can be applied more precisely. For instance, we consider the case where $t = n/3$:

\begin{proposition}\label{proposition:walsh_t_over_3}
    Let $n\in\N^*$ be divisible by $6$. 
    Then, for every $a\in\F_2^n$ and every $k\in[0,n]$, it holds that:
    \[
        |\wtk{d_{n/3,n}}{k}(a)|\leq2^{3n/4},\quad \text{and }\quad|\wt{d_{n/3,n}}(a)|\leq 1+n\cdot 2^{3n/4}.
    \]
\end{proposition}

\begin{proof}
    As before, we want to bound the maximum of $|P_{a,t}(z)|$ on $\dis$ for $a\in\F_2^n$, and this will then be an upper bound on $|\wtk{d_{t,n}}{k}(a)|$ for every $k\in[1,n]$. 
    We write $a=(b,c)$ with $b\in\F_2^{2t}$ and define $(p,q,r)=(p(b),q(b),r(b))$, and further consider $u=n-2t-\w(c)$ and $v=\w(c)$ as in Definition~\ref{defi:P_a_t}.

   To find a bound for $|P_{a,t}(z)|$ on $\dis$ that applies to all $a\in\F_2^n$, we focus only on values of $a$ which are candidates for maximizing $|P_{a,t}(z)|$. 
   Specifically, we consider cases where only one of $(p,q,r)$ is non-zero, and for which only and only one of $(u,v)$ is non-zero.

\pmnote{What is the argument: "maximum implies only one is non-zero" }

   This leaves us with the following cases to consider for $(p,q,r;u,v)$:
    \begin{gather*}
        (t,0,0;n-2t,0),\quad(0,t,0;n-2t,0),\quad(0,0,t;n-2t,0),\\
         (t,0,0;0,n-2t),\quad(0,t,0;0,n-2t),\quad(0,0,t;0,n-2t).
    \end{gather*}
    Since swapping both $(p,q)$ and $(u,v)$ has the same effect as replacing $z$ by $-z$ in $P_{a,t}(z)$ and therefore does not change the maximum of $|P_{a,t}(z)|$ on $\dis$, we can further restrict ourselves to the following tuples for $(p,q,r;u,v)$:
    \[
        (t,0,0;n-2t,0),\quad(0,t,0;n-2t,0),\quad(0,0,t;n-2t,0).
    \]
    Let us for instance find an upper bound of $|P_{a,t}(z)|$ in case $(p,q,r;u,v)=(t,0,0;n-2t,0)$. Since $n-2t=t$ by assumption, we get:
    \[
        P_{a,t}(z)=[(-z^2+2z+1)(1+z)]^t.
    \]
    Writing $z=\cos\alpha+\mathrm i\sin\alpha$, we obtain:

    \pmnote{@Tim, add intermediate computation in the following:}
    \[
        |P_{a,t}(z)|^2=[(6-2\cos2\alpha)(2+2\cos\alpha)]^t.
    \]
    \pmnote{@Tim, for the following, by calculus, you mean using some software? In this case we can add a footnote saying in practice we did it with xxx.

    Can you give a plot of $(\alpha,(6-2\cos2\alpha)(2+2\cos\alpha))$ and add the line corresponding to $y=9/4$?
    }

    Through calculus, one finds that $|(6-2\cos2\alpha)(2+2\cos\alpha)|<2^{9/2}$, and hence that $|P_{a,t}(z)|<2^{9t/4}=2^{3n/4}$. The same upper bound works for the remaining two candidates for $(p,q,r;u,v)$, and we can therefore conclude  $|\wtk{d_{n/3,n}}{k}(a)|\leq 2^{3n/4}$, and $|\wt{d_{n/3,n}}(a)|\leq 1+n\cdot 2^{3n/4}$.
\end{proof}

\pmnote{for the remark, add how they are obtained (from a script in xxx). 
We can give the scripts in supplementary material when submitting, of add the link of a github if you agree to have them public.

\begin{remark}
    The proof of Proposition~\ref{proposition:walsh_t_over_3} can be extended to other values of $t$. 
    In Figure~\ref{fig:walsh_bounds_generalization}, we provide bounds of the form $\max_{a\in\F_2^n}|\wt{d_{\alpha n/40,n}}(a)|\leq 1+n\cdot 2^{\beta n}$, where $\alpha\in[0,20]$ is represented on the $x$-axis and $\beta$ is represented on the $y$-axis. 
\end{remark}

\begin{figure}
	\centering
	\includegraphics[width=9cm]{walsh_bounds_generalization.pdf}
	\caption{Improvements on Corollary~\ref{cor:gen}. For each point $(\alpha,\beta)$, we have $\max_{a\in\F_2^n}|\wt{d_{\alpha n/40,n}}(a)|\leq 1+n\cdot 2^{\beta n}$.}
    \label{fig:walsh_bounds_generalization}
\end{figure}
